---
title: "beacr"
output: 
  rmarkdown::html_vignette:
    toc: false
vignette: >
  %\VignetteIndexEntry{beacr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
tmp_dir <- file.path(tempdir(), "PROJ_ROOT")
dir.create(tmp_dir)
withr::defer(unlink(tmp_dir, recursive = TRUE))
knitr::opts_knit$set(root.dir = tmp_dir)

temp_beacr_dir = fs::path(tmp_dir, "beacr")
dir.create(temp_beacr_dir)

withr::local_options(
  beacr.data_dir_location = temp_beacr_dir,
  beacr.log_dir_location = temp_beacr_dir
)

testthat::local_mocked_bindings(
  .prompt_menu = function(choices, title) {
    return(c(1L))
  },
  .package = 'beacr'
)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 80)
)
```

```{r, include=FALSE}
# Setup to ensure first call is 'fresh', i.e. no existing data dir, no existing log
library(beacr)

reset_seed_log(yes_to_all = TRUE)

# if (fs::dir_exists(beacr:::.get_option("data_dir_location"))) {
#   beacr:::.reset_data_dir(yes_to_all = TRUE)
# }
# beacr:::.reset_data_dir(yes_to_all = TRUE)
#!_RETURN
```

The most simple use case for `beacr` is when you want to set your random number generator seed to a number that is both completely random and reproducible. That code may look something like this:
```{.r}
library(beacr)
seed = beacr::get_seed()
set.seed(seed)
# Continue with random reproducible random processes...
```

When `get_seed()` is called, a seed log is written that captures the details of the seed that was returned based on the beacon(s) used.

```{r}
library(beacr)

# Acquire seed
seed = beacr::get_seed()

# Set seed acquired
set.seed(seed)

# Use reproducible seed
rnorm(10)
```


Then when you rerun the same code, `beacr` will read from the seed log to reproduce the same output.

```{r}
# Acquire seed (automatically reproduces previous seed acquired)
seed = beacr::get_seed()

# Set reproduced seed
set.seed(seed)

# Use reproducible seed
rnorm(10)
```

This log-based workflow makes it possible for an analysis to be completely reproducible without requiring changing of any hard-coded file paths or timestamp values. Note that `get_seed` does not merely _cache_ the Beacon pulse and return the cached value on the second call. Instead, it stores all the information needed to verify and reproduce the pulse used in a `seed_log.json` file. When called for the second time, it re-performs the pulse acquisition and **verifies** that the pulse returned by the Beacon is the same as that logged. Otherwise, users could manually adjust the cached file to make it return any seed value they desired before posting their repository. 

<details>
<summary> Note on what I mean by 'log' </summary>
The **seed log** is a JSON file holding the information needed to reproduce the most recent seed generated from a Beacon. This is not to be confused with the `beacr` **debug log** which is a typical `.log` file retracing all messages, warnings, and errors that occurred during code execution and can be viewed by calling `read_logs()`
</details>

## Pre-registration
If you wish to provide a link to an external file documenting a preregistration of an experimental protocol or hypothesized outcome, currently `beacr` supports using OSF API file IDs. You can pass that like so:
```{.r}
# Publish a preregistration first on OSF that specifies your protocol
# and hypothesized outcomes
#
# ! Note that this file isn't a preregistration but serves as a demo !
OSF_preregistration_file_link = "https://osf.io/mbcw5/files/yahfc"

# Acquire an ex post facto verifiably reproducible seed that is
# inseparably linked to your preregistration
seed = beacr::get_seed(
  preregistration_identifier = OSF_preregistration_file_link,
  preregistration_source = "OSF"
)

# Set reproducible seed
set.seed(seed)

# Use reproducible seed
rnorm(10)
```

When passing an external preregistration, the seed generated by `get_seed(...)` will be uniquely associated with the content of the pre-registration file. In theory, this means that the results of the analysis can only be reproducible if posted alongside the original analysis plan in the form of its preregistration.^1^


<details>
<summary> Super technical specifics about how preregistration is incorporated into seed generation </summary> 
Specifically, the seed value returned by `get_seed` will be (by default, using the NISTBeacon) the SHA-512 hash of the **concatenation** of 

1. The SHA-256 hash of the OSF registration **file data** (NOT the URL)
2. The 512-bit outputValue field of the NISTBeacon JSON API response for the most recent pulse

As a formula, 

`seed = HASH_512( HASH_256(preregistration_binary_data) || BEACON_512 )`

_where `||` means concatenation, not bitwise `OR`_

Before computing this unique seed, `beacr` will verify file integrity by downloading the pre-registration file from the OSF API URI provided, and confirming the hash of the downloaded file matches the hash value provided by the API (i.e. confirming there is no corruption/tampering during transit).

</details>

#### Using a Preregistration
To use a preregistration, first ensure that you've cleared the seed log that you may have already written in prior testing like so:

```{r}
reset_seed_log(yes_to_all = TRUE) 
```


Then continue by passing a preregistration identifier and source. Currently, only OSF files are supported.
```{r}
# Publish a preregistration first on OSF that specifies your protocol
# and hypothesized outcomes
#
# Note that this is not a preregistration but a demo file (actually a 
# talk I gave at PYMS Summer Symposium 2025)
OSF_preregistration_file_link = "https://osf.io/mbcw5/files/yahfc"

# Acquire an ex post facto verifiably reproducible seed that is
# inseparably linked to your preregistration
seed = beacr::get_seed(
  preregistration_identifier = OSF_preregistration_file_link,
  preregistration_source = "OSF"
)

# Set reproducible seed
set.seed(seed)

# Use reproducible seed
rnorm(10)
```

Every subsequent call to `get_seed` will reproduce the Beacon pulse with all the validation steps mentioned above as well as verifying the preregistration file integrity (making sure the file available on OSF hasn't changed since the pulse was originally acquired and logged).

```{r}
# Rerunning the above code...

# Publish a preregistration first on OSF that specifies your protocol
# and hypothesized outcomes
#
# Note that this is not a preregistration but a demo file (actually a 
# talk I gave at PYMS Summer Symposium 2025)
OSF_preregistration_file_link = "https://osf.io/mbcw5/files/yahfc"

# Acquire an ex post facto verifiably reproducible seed that is
# inseparably linked to your preregistration
seed = beacr::get_seed(
  preregistration_identifier = OSF_preregistration_file_link,
  preregistration_source = "OSF"
)

# Set reproducible seed
set.seed(seed)

# Use reproducible seed
rnorm(10)
```

<br>
<br>
<br>
<br>

#### Notes
1. This is not technically true, as currently I use a version of the preregistration file that's condensed to 32 bytes (specifically via the SHA-256bit hash function), so as long as that value is provided, the reproducible seed can be set. This is actually a crucial distinction for future versions of `beacr` which will allow posting _encrypted_ preregistrations that establish primacy but don't reveal what the protocol/hypothesis is until the authors are prepared to release the decryption key. If the authors provide the hash of the secret preregistration then others can reproduce the analysis even **before** they can read the preregistration.
